---
title: "Introduction to ssrfr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ssrfr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ssrfr)
```

`ssrfr` is under development. It guards a single hop: `ssrf_prepare_hop()`
decides whether a URL may be fetched and `ssrf_fetch()` fetches it through a
pinned connection. Following a redirect through the guard comes next. The
contract it implements is `design/specs/ssrfr-v1.md` in the source
repository.

Every hop is evaluated under a policy. It is validated when it is built, so a
malformed rule is an error rather than a rule that silently matches nothing:

```{r}
policy <- ssrf_policy(allow_ranges = "10.0.0.0/8", deny_hosts = ".corp")
policy

try(ssrf_policy(deny_hosts = "com, ru"))
```

The reason codes a refusal will carry are a closed, versioned vocabulary:

```{r}
ssrf_vocabulary("reason_codes")$code
```

## Inspecting a URL

`ssrf_inspect_url()` reports what a URL is under a policy, with no network
I/O. It parses the URL as the guarded fetch will, with `rurl` and then with
libcurl's own parser, and reports the first check that applies: a parse the two
parsers disagree on, the scheme, embedded credentials, the port, an ambiguous
numeric host, the hostname rules, and, for a host written as an address, the
address rules.

```{r}
ssrf_inspect_url("http://0177.0.0.1/")$code
ssrf_inspect_url("http://[64:ff9b::a9fe:a9fe]/latest/meta-data/")
```

It is a pre-filter and a way to lint a configuration, not a defense. It
resolves no name, so a name it finds nothing wrong with may still point
anywhere:

```{r}
ssrf_inspect_url("https://example.com/")$code

policy <- ssrf_policy(allow_ranges = "10.0.0.0/8")
ssrf_inspect_url("http://10.1.2.3/", policy)$code
```

The provider endpoints and metadata hostnames it refuses by default are
vocabularies too, each row citing the vendor documentation that names it:

```{r}
ssrf_vocabulary("metadata_hostnames")$hostname
```

## Inspecting what a name resolves to

With `layer = "L1"`, `ssrf_inspect_url()` also asks the system resolver, once,
for every address of a name the earlier checks passed, and classifies each
address with the addresses embedded in it. The name is looked up as absolute,
with a trailing root dot, so `http://intranet/` is never completed with a DNS
search domain. One refused address refuses the whole set, and each address
keeps its own facts. Were `dual.example` to resolve to `93.184.216.34` and
`::1`:

```{r, eval = FALSE}
res <- ssrf_inspect_url("http://dual.example/", layer = "L1")
res$code
#> [1] "loopback"
vapply(res$addresses, function(a) a$code, "")
#> [1] NA         "loopback"
```

A resolver error, an empty answer or an answer that is not an address is the
operational cause `"unresolvable"`, in `cause` rather than `code`. A URL an
earlier check refuses is never looked up, so this makes no DNS query:

```{r}
ssrf_inspect_url("http://files.example:6379/", layer = "L1")$code
```

L1 is still not a defense: the answer can change the moment it returns. Only
the guarded fetch, which connects to an address it has just validated, is.

## Fetching through the guard

A guarded hop has two steps. `ssrf_prepare_hop()` parses the URL, checks it,
resolves a name once and classifies every address, and returns one of three
values: a refusal (`code`), an operational failure (`cause`), or a binding.
Only a binding can be fetched. Nothing below makes a network call: these URLs
are refused, or their host is an address, which needs no resolution.

```{r}
policy <- ssrf_policy()

refusal <- ssrf_prepare_hop("http://169.254.169.254/latest/", policy,
                            request = list())
refusal$code
ssrf_public_reason(refusal)

binding <- ssrf_prepare_hop(
  "https://93.184.216.34/data?page=2#top",
  policy,
  request = list(headers = c(Authorization = "Bearer not-shown"))
)
binding
```

The request plan is part of the decision, not of the fetch: a list of
`method`, `headers`, `body` and `carry`, where `list()` is a plain `GET`. A
header the transport owns, such as `Host`, or a metadata-service marker such
as `Metadata-Flavor`, is an error before anything is parsed:

```{r}
try(ssrf_prepare_hop("https://example.com/", policy,
                     request = list(headers = c(`Metadata-Flavor` = "Google"))))
```

`ssrf_fetch()` then connects only to an address the binding validated. The
connection is pinned with libcurl's `connect_to`, so the `Host` header, SNI
and certificate verification stay bound to the hostname, while the name is
never resolved again. The binding is single-use and spent on entry; the
response owns no connection and prints no body:

```{r, eval = FALSE}
response <- ssrf_fetch(binding)
response
#> <ssrfr_response>
#>   status: 200
#>   type: text/html
#>   body: 1256 bytes (read it with $body)
rawToChar(response$body)
binding$state$status
#> [1] 200
try(ssrf_fetch(binding))
#> Error : This binding was already fetched; prepare the hop again.
```

A redirect is returned, not followed: its status is in the response and its
`Location` in `binding$state$location`. A failure on the wire is a value, not
an error: `cause` is one of `ssrf_vocabulary("causes")`, and `detail` lists
the addresses tried and how each attempt ended. Like a refusal, project it
with `ssrf_public_reason()` before an untrusted party sees it.

## What the guard cannot protect

`ssrfr` guards only requests made through a binding. These stay unguarded,
and an application that lets an attacker influence the URL must not use them
for it:

- base R's `download.file()`, `url()`, and `readLines()`, `read.csv()` or
  `readRDS()` given a URL;
- direct `curl` calls (`curl::curl_fetch_memory()`, `curl::curl()` and the
  rest), and the `curl` command-line tool, whether through `system()` or
  `download.file(method = "curl")`, which also reads `~/.curlrc`;
- the packages that read URLs through them, such as
  `jsonlite::fromJSON(url)`, `data.table::fread(url)`, `xml2::read_xml(url)`
  and `xml2::read_html(url)`.

`ssrfr` is not egress control either: a firewall or a separate network is
the stronger boundary, and `ssrfr` is defense in depth beside it.

## Testing against a local server

A loopback server is itself a refused destination, so a test that fetches
one through the guard reopens loopback in its own policy. There is no test
mode and no switch that turns the guard off; keep this policy in tests:

```{r, eval = FALSE}
test_policy <- ssrf_policy(allow_ranges = "127.0.0.0/8", allow_ports = 8080)
binding <- ssrf_prepare_hop("http://127.0.0.1:8080/health", test_policy,
                            request = list())
ssrf_fetch(binding)$status
```
