---
title: "Introduction to ssrfr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ssrfr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ssrfr)
```

`ssrfr` is under development and does not guard a fetch yet. The contract it
implements is `design/specs/ssrfr-v1.md` in the source repository. This
vignette will show a guarded fetch and a caller-owned redirect loop once
`ssrf_prepare_hop()` and `ssrf_fetch()` exist.

What exists so far is the policy a hop will be evaluated under. It is
validated when it is built, so a malformed rule is an error rather than a rule
that silently matches nothing:

```{r}
policy <- ssrf_policy(allow_ranges = "10.0.0.0/8", deny_hosts = ".corp")
policy

try(ssrf_policy(deny_hosts = "com, ru"))
```

The reason codes a refusal will carry are a closed, versioned vocabulary:

```{r}
ssrf_vocabulary("reason_codes")$code
```

## Inspecting a URL

`ssrf_inspect_url()` reports what a URL is under a policy, with no network
I/O. It parses the URL as the guarded fetch will, with `rurl` and then with
libcurl's own parser, and reports the first check that applies: a parse the two
parsers disagree on, the scheme, embedded credentials, the port, an ambiguous
numeric host, the hostname rules, and, for a host written as an address, the
address rules.

```{r}
ssrf_inspect_url("http://0177.0.0.1/")$code
ssrf_inspect_url("http://[64:ff9b::a9fe:a9fe]/latest/meta-data/")
```

It is a pre-filter and a way to lint a configuration, not a defense. It
resolves no name, so a name it finds nothing wrong with may still point
anywhere:

```{r}
ssrf_inspect_url("https://example.com/")$code

policy <- ssrf_policy(allow_ranges = "10.0.0.0/8")
ssrf_inspect_url("http://10.1.2.3/", policy)$code
```

The provider endpoints and metadata hostnames it refuses by default are
vocabularies too, each row citing the vendor documentation that names it:

```{r}
ssrf_vocabulary("metadata_hostnames")$hostname
```
