---
title: "Introduction to ssrfr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ssrfr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ssrfr)
```

`ssrfr` is in its design phase and exports no guard functions yet. The
contract it will implement is `design/specs/ssrfr-v1.md` in the source
repository. This vignette will show a guarded fetch and a caller-owned redirect
loop once `ssrf_prepare_hop()` and `ssrf_fetch()` exist.

What exists so far is the policy a hop will be evaluated under. It is
validated when it is built, so a malformed rule is an error rather than a rule
that silently matches nothing:

```{r}
policy <- ssrf_policy(allow_ranges = "10.0.0.0/8", deny_hosts = ".corp")
policy

try(ssrf_policy(deny_hosts = "com, ru"))
```

The reason codes a refusal will carry are a closed, versioned vocabulary:

```{r}
ssrf_vocabulary("reason_codes")$code
```
